### 2018-07-06 paper3_relevance and symbols_01.Rmd
Add files via upload

---
title: "The persuasion of symbolic elements in government public communications, and their moderation by the personal relevance of the communication"
author: '**Saar Alon-Barkat**, PhD candidate; the Hebrew University of Jerusalem;   saar.barkat@mail.huji.ac.il'
date: " "
output:
  word_document: default
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      toc_depth: 3
link-citations: yes
bibliography: phd_paper_2.bib
urlcolor: blue
---

<br>

Last edited at `r Sys.Date()`.[^footnote_scripts]

[^footnote_scripts]: Original R scripts available at:  https://github.com/saaralonbarkat/symbols_and_relevance

```{r , include=FALSE, echo=FALSE}
#load("SVIVA_R_ENV.RData")
source("SVIVA_exp2_dm_03.R")

SVIVA2_01_haifa = filter(SVIVA2_01,AREA==1)
SVIVA2_01_center = filter(SVIVA2_01,AREA==0)
```

```{r silent-packages, echo = FALSE, eval = TRUE, message=FALSE, include = FALSE}
library(knitr)
library(tidyverse)
library(stargazer)
library(car)
library(broom)
library(kableExtra)
```

```{r,echo=FALSE,warning=FALSE,message=FALSE}
N_raw = nrow(SVIVA2_raw_00)
N_0=SVIVA2_00%>%nrow()
N_1=SVIVA2_01%>%nrow()

filter_IP = N_0-(SVIVA2_00 %>%
  distinct(IP,.keep_all=TRUE)%>%nrow())
filter_age = N_0-(SVIVA2_00 %>%
  filter(!(AGE%in% 1:17))%>% nrow())
filter_IMC = N_0-(SVIVA2_00 %>%
  filter(IMC==1)%>% nrow())
filter_time = N_0-(SVIVA2_00 %>%
  filter(TIMER_total>=3,TIMER_total<=30)%>%nrow())

N_1_haifa = filter(SVIVA2_01,AREA==1)%>%nrow()
N_1_center = filter(SVIVA2_01,AREA==0)%>%nrow()

```

<br>

---

# Abstract

*Recent experimental studies have demonstrated that government public communications can enhance citizens' trust in government organizations and their policies not only by conveying substantive information, but also by incorporating strategically desigened symbolic element that evoke positive feelings and associations. Yet, under which conditions are citizens more/less likely to be persuaded by symbols? Building on the social psychology Elaboration Likelihood Model (ELM), I theorize that citizens are less susceptible to symbols and more likely to scrutinize the substantive content of the communication, the more they perceive the policy issue in the communication as personally relevant. I test this theory through a survey experiment, where I examine Israeli citizens' trust in the Environment Protection Ministry's policy plans for reducing the air pollution in a specific area (the Haifa-Bay). The perceived personal relevance of the policy is operationalized both experimentally, and by comparing between citizens from the polluted Haifa-Bay area and others. Counter to the hypotheses, these comparisons suggest that higher perceived personal relevance does not attentuate the effect of symbols. The theoretical and norative implications of these findings are discussed.*

---

<br>

# Introduction

A growing public administration literature explores whether government public communications can enhance citizens' trust in public organizations and policies. Whereas most of the studies have focused on the *informative* aspects of government communications [e.g. @marvel_2015; @im_2014; @porumbescu_2017; @porumbescu_2016; @grimmelikhuijsen_2015; @ho_2017], recent experimental studies further pointed to the emotive effect of *symbolic* elements entangled in them, such as  brand names [@teodoro_2018], logos [@karens_2016], and figures [@alonbarkat_2017]. Such symbols and others exist in almost every government communication, and most prominently in public campaigns, which have been increasingly used by public-sector organizations [@eshuis_2012; @marland_2017].

While previous studies have substantiated the distinct effect of symbols on citizens' trust, they did not theorize, nor they empirically examined the conditions under which are these symbols are likely to be more/less effective. This research gap has great significance. First, from the practical point of view of government organizations, this may enable them to design more effective communications which may enhance citizens' collaboration with their policies. Second, employing a broader perspective of democratic theory, one can argue that the emotive effect of symbols challenges the fundamental normative expectation that citizens' judgments should be based on their critical thinking. Furthermore, the positive symbolic message can compensate for poor performance and create a misalignment between citizens' perceptions and the "real world" [@alonbarkat_2017]. Hence, by identifying the circumstances under which citizens are more/less likely to be affected by symbols, we would better assess the potential risk for governments' manipulative use of public communications.

To address this literature gap, I draw upon the social psychology Elaboration Likelihood Model (ELM) [@petty_1986; @petty_2011]. I hypothesize that citizens are less susceptible to symbols the more they perceive the policy issue in the communication as having personal relevance for themselves, which enhances their motivation to invest in elaborate processing. Similarly, higher perceived personal relevance are expected to enhance citizens' inclination to scrutinize the substantive content of communication. Employing a survey experiment, I test this hypothesis in the context of the Israeli Environment Protection Ministry, and its policy plan for reducing the air pollution in a specific area (the Haifa-Bay). I operationalize the perceived personal relevance of the policy plan both experimentally, and by comparing between citizens from the polluted Haifa-Bay area and others.

<br>

# The Elaboration Likelihood Model (ELM) and the effect of symbols and information in communicaions.

The Elaboration Likelihood Model (ELM), developed in the late 1970s by social psychology scholars Richard Petty and John Cacioppo, provides a theoretical framework for investigating the effects of "persuasive" communications on people's attitudes [@petty_1986; @petty_2011]. According to ELM, people may respond differently to a communication, depending on the amount of thinking that they dedicate to it. When people have high motivation and ability to elaborate on the message, they are likely to focus on the arguments in the communication which are central to the merits of the issue, and scrutinize them ("central route"). Conversely, when they are unmotivated or unable to elaborate, they may rely on available heuristic cues that require relatively little amount of cognitive effort ("peripheral route"). These two routes for persuasion are regarded as two ends of an "elaboration continuum". People can be also located between these two edges, and then persuasion will be determined by a mixture of these processes.

In light of the latter conceptualization of ELM, substantive information in government public communications represent message arguments, which can be critically evaluated based on their logical quality (central route). Conversely, symbolic elements can be regarded as potential peripheral cues, which can affect attitudes by evoking emotions and associations attributed to government organizations and their policies (peripheral route). Accordingly, the more citizens are motivated to think about the content of the communication, and the more they are capable of doing so, the more they are likely to form their attitudes based on its substantive information. Conversely, the less motivated and capable they are to elaborate, the more susceptible they are to symbolic elements in the communication.

This literature has pointed to a variety of factors that shape people's motivation and ability to think about the message, and accordingly affect their inclination to be persuaded either by the arguments or by peripheral cues. Among others, scholars have theorized that the more people perceive the issue as personally relevant, the more motivated they are to invest in thinking about it, since the consequences of being incorrect are greater. That postulation has been empirically tested, and largely confirmed, by numerous ELM based experimental studies in social psychology and marketing [@petty_1979; @petty_1981; @petty_1983; @johnson_1989; @petty_1990; @burnkrant_1989]. These studies have demonstrated that increasing the perceived personal relevance of a communication enhances the effect of arguments in the communication and attenuates the effect of peripheral cues. Later reviews of ELM literature considered the factor of perceived personal relevance as the most important determinant of motivation to process a message [@petty_2002; @petty_2011].[^footnote_elm_factors]

[^footnote_elm_factors]: For a thorough review of motivational and ability factors affecting likelihood to elaborate see @petty_1998.

Connecting the above public administration literature on citizens' responses to government public communications with the social psychology ELM literature, I theorize that:

*H<sub>1</sub> - Symbolic elements in communications are less (more) likely to increase citizens' trust in governemnt organizations and policicies, the more (less) citizens perceive the policy issue in the communication as having personal relevance for themselves.*

*H<sub>2</sub> - The logical quality of the communication is more (less) likely to affect citizens' trust in governemnt organizations and policicies, the more (less) citizens perceive the policy issue in the communication as having personal relevance for themselves.*

These theoretical hypotheses are graphically displayed in **Figure 1**.

<br>
**Figure 1: Theoretical model**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/theoretical_model_relevance.png)

<br>

# Research design

I tested the above mentioned hypotheses via a randomized survey experiment, building on the methodology of @petty_1983. I focus on the case of the Israeli Environment Protection Ministry (hereafter: EPM), and examines citizens' responses to communications of EPM regarding two of its current policy plans. The first policy regards *reducing the air pollution in the Haifa-Bay area*, located in the north of Israel. This policy issue offers a fortunate case for examining the moderating role of perceived personal relevance. The differences between those who reside in the polluted area and those who do not represent natural variation in this factor. Moreover, because this policy issue is intangible, ambiguous and complicated, the differences in personal relevance are not strongly confounded with differences in personal knowledge, as it is the case in many other policy issues [@petty_1983]. Congruently, those who reside in the polluted Haifa-Bay area are more motivated to scrutinize a communication about the air-pollution in their area of residence, yet are not more capable to do so than those who reside elsewhere. As explained below, the sample population of the study consists of residents of Haifa-bay area, matched with residents of cities at the center of Israel with similar aggregate socio-economic profile. The second policy plan regards *reducing domestic waste and increasing recycling*. This topic is assumed to be equally relevant for Haifa-bay residents and others. Therefore, adding this policy plan enables to further confirm that their different responses to the air-pollution policy, if found, are explained by differences in perceived personal relevance. I henceforth refer to these two policy plans as "air-pollution" and "recycling".

I focused on three symbolic elements which have been used in the EPM's public communications, and became widely recognized by the public. The first symbolic element is the unique brand logo of the EPM, which consists a pair of green and orange leaves that resemble two hands, designed to symbolize peace and harmony with the environment. The second element is the green color, which is strongly associated with taking care of the environment. The third element is the images of two high-profile celebrity comedians which were taken from advertising campaigns of the EPM. Further details about these symbolic elements are found in the supplementary appendix.

<br>

## *Data collection*

The survey was conducted online through Qualtrics Survey Software. To capture natural variation in the personal relevance of the air-pollution policy, I sampled citizens from the major cities in the Haifa-Bay area, and matched them with citizens from cities with a similar profile and population size at the center of the country (hereafter "center").[^footnote_cities] To select and recruit residents of the selected cities, I was assisted by an Israeli internet research panel company *iPanel*. The online survey link was sent by iPanel, between 28.1-1.2.2018. A total of `r N_0` respondents completed the survey (a response rate of `r round((N_0/5077)*100,1)`%). After filtering,[^footnote_filters] I ended up with a sizable sample of `r N_1` subjects, of whom `r SVIVA2_01_haifa%>%nrow()` live in the Haifa-bay cities, and `r SVIVA2_01_center%>%nrow()` live in the center cities.

[^footnote_cities]: The cities from the Haifa-Bay area are: *Haifa*, *Nesher*, *Qiryat-Hayiim*, *Qiryat-Bialik*, *Qiryat-Mozkin*, *Qiryat-Ata* and *Qiryat-Yam*. The matched cities from the center are: *Natania*, *Petach-Tikva* and *Rishon-Lezion*. The socioeconomic characteristics of the cities are summarized in the supplementary appendix.

[^footnote_filters]: I filtered out `r N_0-N_1` observations (`r round(((N_0-N_1)/N_0)*100,1)`%) due to multiple entries from the same IP address (n=`r filter_IP`), failure at the instructional manipulation test (n=`r filter_IMC`), submitting the survey in less than 3 or more than 30 minutes (n=`r filter_time`), and respondents under the age of 18 (n=`r filter_age`). These filters are niether associated with the areas nor with the experimental conditions.

<br>

## *Procedure*

The survey was presented to the participants as a study about citizens' attitudes regarding environmental policies in general. After a set of pre-manipulation questions, they subjects were randomly assigned to treatment and control conditions of perceived personal relevance (hereafter: relevance). Subjects in the relevance treatment group were asked three questions about their interest in environmental policy issues, the extent to which they follow these issues in the media, and their preference for more/less government investment in this field. These questions were designed to prime subjects to think about the impact of environmental issues to their own lives. Subjects were also asked to describe the most salient environmental issue in their particular area of residence. The latter open question was specifically designed to prime Haifa-bay residents to think about the personal relevance of the air pollution in their area. Subjects in the relevance control group were asked equivalent questions about their field of occupation.

Participants were then presented with the two policy plans of EPM, which they were told that were obtained from EPM's annual work plan. Each policy plan was followed by questions about respondents' trust in them. Per each policy, subjects were randomly allocated to one of two conditions of substantive information (hereafter: information): strong policy (i.e. logically persuasive policy plan) or weak policy (i.e. logically unpersuasive). The information in the strong policy plans condition was taken from the real work plan and from other official sources, whereas the information in the weak policy plans was fictional, and was deliberately designed to create a poor policy plan, which is incompatible with achieving the policy goals. The full texts of the weak and strong conditions for the two policy plans is displayed in **Appendix A**. Each subject, moreover, was exposed to one strong and one weak policy communication, displayed in a random order.

The participants were also randomly assigned to one of three conditions of symbolic elements in the communication (hereafter: symbols): treatment ("real symbols"), control-"no symbols" and control-"fake symbols". In the symbol treatment condition, the two policies (again, one weak and one strong) are displayed with the EPM logo, green color and celebrity images from campaigns. In the control no-symbols condition, the two policies are displayed in a minimal black and white design, and without the logo and the celebrities images. In the control fake-symbols condition, the EPM logo is replaced by a fake logo, the green color is replaced with blue, and the images of celebrities from the campaigns are replaced with images of unfamiliar people. The symbols conditions are displayed in **Appendix B**. Further details about the real and fake symbols are found in the supplementary appendix.

Altogether, per each policy plan, the experiment has a factorial design of 3 (symbols: real symbols, fake-symbols, no-symbols) x 2 (information: strong vs. weak policy) x 2 (relevance: treatment vs. control).[^footnote_information_symbols_paper]  The assignment to the different experimental groups, across the two areas, is graphically presented in **Figure 2**. The experimental groups are fairly balanced with respect to the two areas, as well as to demographics, and pre-manipulation variables.

[^footnote_information_symbols_paper]: In addition to the above-mentioned hypotheses, the experiment was also designed to test a different set of hypotheses regarding the interaction between symbols and information, and its moderation by the effect of symbols on elaboration. In this paper, however, I focus on those methodological and empirical parts of the study that are consequential for the examination of the moderating role or perceived personal relevance. The methodological aspects which are relevant for examining the additional hypotheses are described in a separate paper.

<br>
**Figure 2: Experimental design for policy plan communications**
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/experiment_flow_relevance.png)

<br>

##*Variables*

```{r, echo=FALSE, warning=FALSE, message=F}
alpha_air_trust = psych::alpha(select(SVIVA2_01, TRUST_air_q1,
                               TRUST_air_q2,
                               TRUST_air_q3,
                               TRUST_air_q4,
                               TRUST_air_q5,
                               TRUST_air_q6))[["total"]][["raw_alpha"]]%>%round(2)

alpha_waste_trust = psych::alpha(select(SVIVA2_01, TRUST_waste_q1,
                               TRUST_waste_q2,
                               TRUST_waste_q3,
                               TRUST_waste_q4,
                               TRUST_waste_q5,
                               TRUST_waste_q6))[["total"]][["raw_alpha"]]%>%round(2)

alpha_trust_comb = ((alpha_air_trust+alpha_waste_trust)/2)%>%round(2)

```

My outcome variable, citizens' trust in policy, is measured in the survey using a composite index of six items (Cronbach’s alpha = `r alpha_air_trust` and `r alpha_waste_trust` for the air-pollution and waste policies respectively). This scale was adopted with modifications from validated trust scales used by previous public-administration studies [@karens_2016; @grimmelikhuijsen_2017; @grimmelikhuijsen_2018]. Participants were asked to indicate their agreement with the following statements, between 1 (weakly agree) and 7 (strongly agree): (1) I believe that the actions mentioned in the policy plan will assist in fulfilling the policy goal; (2) I believe that the actions mentioned in the policy plan were designed in a professional manner; (3) I believe that the policy plan is in the interest of citizens; (4) I believe that the policy plan reflects a genuine attempt to improve the well-being of citizens; (5) I believe that EPM made an honest attempt to design a good policy plan; (6) I believe that the Ministry of Environmental Protection aims to keep its commitments in that policy plan. Participants reported lower trust, on average, in the air-pollution policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max() %>% round(2)`) compared with the recycling policy (*Mean* = `r SVIVA2_01_comb%>%summarise(mean(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`, *SD* = `r SVIVA2_01_comb%>%summarise(sd(TRUST_waste_INDEX,na.rm=T))%>%max() %>% round(2)`).

My key independent variables are the manipulations of symbols and information in the communication, as described above, and their interactions with the perceived personal relevance of the communication.[^footnote_symbols_manipulation_check] The perceived personal relevance of the communication is operationalized observationally (through the distinction between the two areas: Haifa-Bay and center) as well as experimentally (through the abovementioned relevance manipulation). I coded living in Haifa-bay area according to respondents’ reporting on residence, work or studying in one of the Haifa-Bay cities. My data confirmes that the air-pollution policy is perceived as more personally relevant by Haifa Bay residents compared with center residents, whereas the recycling policy is perceived by both groups as equally relevant. Moreover, when directly asked about the most salient environmental issue in their area of residence, most of the Haifa-Bay subjects regarded to the air-pollution. As for the relevance priming questions, however, manipulation checks suggest that those in the relevance treatment group did not perceive either of the two policies as more personally relevant. This limitation should be kept in mind when interpreting the results regarding this manipulation. The results of the all manipulation and validation checks are available from the supplementary appendix, along with summary statistics and correlation matrix for all variables.

[^footnote_symbols_manipulation_check]: Manipulation checks for the symbols confirm that the subjects in both areas are fairly familiar with the EPM symbols, and these symbols evoke positive feelings.

<br>

# Results

```{r,results="asis",echo=FALSE,warning=FALSE,message=FALSE}
sd_trust_air = SVIVA2_01_comb%>%summarise(sd(TRUST_air_INDEX,na.rm=T))%>%max()
sd_trust_waste = SVIVA2_01_comb%>%summarise(sd(TRUST_waste_INDEX,na.rm=T))%>%max()

mod_air_1.1 = lm(TRUST_air_INDEX~SYMBOL_t+INFORMATION_weak,data=SVIVA2_01_comb)
mod_air_1.2 = update(mod_air_1.1,. ~ .+AREA_center)
mod_air_1.3 = update(mod_air_1.1,. ~ .+AREA_center*(SYMBOL_t+INFORMATION_weak))

mod_waste_1.1 = lm(TRUST_waste_INDEX~SYMBOL_t+INFORMATION_weak,data=SVIVA2_01_comb)
mod_waste_1.2 = update(mod_waste_1.1,. ~ .+AREA_center)
mod_waste_1.3 = update(mod_waste_1.1,. ~ .+AREA_center*(SYMBOL_t+INFORMATION_weak))

mod_air_1.1_r1 = lm(TRUST_air_INDEX~SYMBOL_t.r+INFORMATION_weak,data=SVIVA2_01_comb)
mod_waste_1.1_r1 = lm(TRUST_waste_INDEX~SYMBOL_t.r+INFORMATION_weak,data=SVIVA2_01_comb)

```

**Table 1** presents the results of OLS regression analyses for the effects of the symbols and information manipulations on trust in the two policy plans, and their interaction with the two areas. Models 1.1-1.3 regard the Air-pollution policy, and Models 1.4 - 1.6 regard the recycling policy. I regress trust in policy on the manipulations (Models 1.1, 1.4), then I add the areas (Models 1.2, 1.5) and the interaction between them (Models 1.3, 1.6). The real symbols had positive main effects on both policies compared with the no symbols condition (equivalent to `r ((mod_air_1.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())/sd_trust_air)%>%round(2)%>%abs()` and `r ((mod_waste_1.1%>%tidy()%>%filter(term=="SYMBOL_t1")%>%select(estimate)%>%max())/sd_trust_waste)%>%round(2)%>%abs()` standard deviations for the air-pollution and recycling policies, respectively). Compared with the fake symbols, the increase is significant at the 90% (one-tailed) regarding the air-pollution policy, and only marginally significant compared with the recycling policy (*p* = `r (summary(mod_waste_1.1_r1)$coefficients["SYMBOL_t.r1",4]/2)%>%round(2)`, one-tailed). The information manipulation also has a significant main effect in both policy plans, as expected (equivalent to `r ((mod_air_1.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())/sd_trust_air)%>%round(2)%>%abs()` and `r ((mod_waste_1.1%>%tidy()%>%filter(term=="INFORMATION_weak")%>%select(estimate)%>%max())/sd_trust_waste)%>%round(2)%>%abs()` standard deviations, respectively). Regarding the two areas, as shown by the coefficients of this factor in Models 1.2 and 1.5, participants from Haifa-Bay reported lower levels of trust in the air-pollution policy, and also, to a lower degree, in the recycling policy. The latter effects suggests that Haifa-Bay residents tend to have prior negative attitudes towards EPM's actions regarding the air-pollution in their area, and to some extent towards EPM or environmental policy in general.

Keeping in mind these main effects, I now turn to analyze the interactions between the manipulations and the areas. With regard to the air-pollution policy, my first hypotheses entails a positive interaction between residence in the center (i.e. low personal relevance) and the real symbols. Based on my second hypothesis, I expect to find a positive interaction between center residents and weak information. Counter to these expectations, in Model 1.3, the interaction between residence in the center and the real symbols is negative (*p*<.1), and the interaction with the weak information is insignificant. These findings suggest that the real symbols were more effective in increasing the trust of those reside in the polluted area and that the latter were not more likely to scrutinize the content of the communication. I also found similar patterns regarding the recycling policy. In Model 1.6, the interaction between the areas and the information is insignificant, and the interaction with the real symbols is negative. **Figure 3** provides a descriptive demonstration of the interaction between the areas and the symbols manipulation for the two policies.

**Table 1: Regression analyses – observational measure for perceived personal relevance.**

```{r,results="asis",echo=FALSE,warning=FALSE,message=FALSE}
stargazer(mod_air_1.1,mod_air_1.2,mod_air_1.3,
          mod_waste_1.1,mod_waste_1.2,mod_waste_1.3,
          type="html",style = "apsr",
          dep.var.labels=c("Trust in air-pollution policy","Trust in recycling policy"),
          covariate.labels=c("Real symbols",
                             "Fake-symbols",
                             "Weak Information",
                             "Area (0=Haifa-Bay; 1=center)",
                             "Area x Real symbols",
                             "Area x Fake symbols",
                             "Area x Weak Information"),
          omit.stat = c("rsq","ser", "f"),
          initial.zero = FALSE,
          notes = "*Notes*: Table entries are nonstandardized OLS-regression coefficients, with Standard errors in parentheses. The reference category for the symbols manipulation conditions is the control (no symbols).",
          notes.append = FALSE)

```

&ast;*p*<.1; &ast;&ast;*p*<.05; &ast;&ast;&ast;*p*<.01 (two-tailed).

**Figure 3: The effect of symbolic elements on trust across areas**

```{r, echo=FALSE,warning=FALSE,message=FALSE, fig.width=8,fig.height=4}
SVIVA2_01_comb = SVIVA2_01_comb %>% 
  mutate(POLICY_n = Recode(policy,"'TRUST_air_INDEX'='Air-pollution'; 'TRUST_waste_INDEX'='Recycling'"))

ggplot(SVIVA2_01_comb, aes(factor(SYMBOL_n),trust,shape=factor(AREA_names)))+ 
  theme_classic()+
  #geom_boxplot(width=0.25,color="gray70")+
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar",
               fun.args = list(mult = 1.65),
               width=0.2,size=0.5,position=position_dodge(0.2),color="gray70") + 
  stat_summary(fun.y = mean, geom = "point", size=1.6,position=position_dodge(0.2))+
  stat_summary(fun.y = mean, geom = "line", size=0.6, aes(group = AREA_names,linetype=AREA_names),position=position_dodge(0.2),alpha=0.8)+
  scale_shape_manual(guide=F,values=c(19,1))+
  scale_linetype_manual(name="Area",values=c(1,4))+
  facet_grid(~POLICY_n)+
  scale_x_discrete(name="",labels=c("No symbols","Fake symbols","Real symbols"))+
  scale_y_continuous(name = "Trust in policy",breaks = seq(1,7,by=0.1))

```

<font size="0.5">

*Note*: Error bars represent 90% confidence intervals.

</font>

Next, I examine whether those subjects who were primed to think about the personal relevance of these policy plans were more likely to scrutinize them, and less likely to be affected by the familiar symbols. The interaction between the relevance manipulation and the symbols and information manipulations is presented in **Table 2**. Again, the first three models regard the Air-pollution policy, and the remaining models regard the recycling policy. Models 2.1, 2.4 replicate models 1.1, 1.4 while adding the relevance manipulation. Thereafter, I add the interactions between the relevance manipulation and the symbols and information manipulations (Models 2.2, 2.5). Finally, to test the moderation of the relevance manipulation within each area, I add three-way interactions with the areas (Models 2.3, 2.6). In all these models, the relevance treatment did not yield significant effects, and there are no significant interactions with the symbols and information manipulations. The triple interactions with the areas are also insignificant (in Models 2.3, 2.6), which indicates that in neither of these areas did the relevance treatment moderate the effects of the symbols and information manipulations.

**Table 2: Regression analyses – experimental manipulation of perceived personal relevance.**

```{r,results="asis",echo=FALSE,warning=FALSE,message=FALSE}
mod_air_2.1 = lm(TRUST_air_INDEX~SYMBOL_t+RELEVANCE_exp+INFORMATION_weak,data=SVIVA2_01_comb)
mod_air_2.2 = lm(TRUST_air_INDEX~RELEVANCE_exp*(SYMBOL_t+INFORMATION_weak),data=SVIVA2_01_comb)
mod_air_2.3 = lm(TRUST_air_INDEX~RELEVANCE_exp*(SYMBOL_t+INFORMATION_weak)+SYMBOL_t*AREA_center,data=SVIVA2_01_comb)
mod_air_2.4 = lm(TRUST_air_INDEX~RELEVANCE_exp*(SYMBOL_t+INFORMATION_weak)+SYMBOL_t*AREA_center+RELEVANCE_exp*SYMBOL_t*AREA_center+RELEVANCE_exp*INFORMATION_weak*AREA_center,data=SVIVA2_01_comb)

mod_waste_2.1 = lm(TRUST_waste_INDEX~SYMBOL_t+RELEVANCE_exp+INFORMATION_weak,data=SVIVA2_01_comb)
mod_waste_2.2 = lm(TRUST_waste_INDEX~RELEVANCE_exp*(SYMBOL_t+INFORMATION_weak),data=SVIVA2_01_comb)
mod_waste_2.3 = lm(TRUST_waste_INDEX~RELEVANCE_exp*(SYMBOL_t+INFORMATION_weak)+SYMBOL_t*AREA_center,data=SVIVA2_01_comb)
mod_waste_2.4 = lm(TRUST_waste_INDEX~RELEVANCE_exp*(SYMBOL_t+INFORMATION_weak)+SYMBOL_t*AREA_center+RELEVANCE_exp*SYMBOL_t*AREA_center+RELEVANCE_exp*INFORMATION_weak*AREA_center,data=SVIVA2_01_comb)

stargazer(mod_air_2.1,mod_air_2.2,mod_air_2.4,
          mod_waste_2.1,mod_waste_2.2,mod_waste_2.4,
          type="html",style = "apsr",
          dep.var.labels=c("Trust in air-pollution policy","Trust in recycling policy"),
          covariate.labels=c("Real symbols",
                             "Fake-symbols",
                             "Weak Information",
                             "Relevance treatment",
                             "Relevance treatment x Real symbols",
                             "Relevance treatment x Fake symbols",
                             "Relevance treatment x Weak Information",
                             "Area (0=Haifa-Bay; 1=center)",
                             "Area x Real symbols",
                             "Area x Fake symbols",
                             "Area x Weak Information",
                             "Relevance treatment x Area",
                             "Relevance treatment x Area x Real symbols",
                             "Relevance treatment x Area x Fake symbols",
                             "Relevance treatment x Area x Weak Information"),
          omit.stat = c("rsq","ser", "f"),
          order = c(1,2,4,3,6:8,5,9,10,12,11,13:15),
          initial.zero = FALSE,
          notes = "",
          notes.append = FALSE)

```

<br>

I further explored the robustness of these findings through additional analyses, reported in the supplementary appendix. *First*, in addition to the above operationalizations of personal relevance, I also relied on subjects' subjective reporting. *Second*, I further explored the effect of personal relevance on subjects inclination to scrutinize the policy plans, using additional measures for message elaboration available in the survey: the time participants spent on scanning the policy plans and their score in a memory test. I also tested whether personal relevance decreased the delta (within-subjects) between trust in strong and weak policy plans presented to each subject. *Third*, I added a set of individual-level controls, and factors related to personal relevance of air-pollution, and to personal interest in environmental issues. Finally, I replicated my analyses on the unfiltered sample. None of these multiple robust analyses supports my theoretical hypotheses.

<br>

#Discussion and conclusion

This study contributes to the growing public administration research on citizens' responses to government public communications, and specifically to symbolic elements entangled in these communications. I explored the postulation, derived from social psychology, that citizens are less susceptible to symbols and more likely to scrutinize the substantive information in the communication the more they perceive it as personally relevant. I put this theory to rigorous empirical test, using a randomized survey experiment focusing on the Israeli Environmental Protection Ministry. I further focused on its policy plan regarding air pollution in a specific area, to capture natural variation in citizens' perceptions of the personal relevance. In addition to utilizing these natural differences, I also used manipulated the personal relevance of the policy plans. To the best of my knowledge, this is the first study to assess the  moderating role of perceived personal relevance on citizens' responses to government symbolic communications.

Whilst the effect of perceived personal relevance on message processing has been substantiated by experiment-based studies in social psychology and marketing, I did not find support for this hypothesis the context of citizens' responses to government communications in this study. Subjects residing in the polluted Haifa-Bay indeed reported that they perceive the policy plan for reducing air-pollution in that area as highly relevant for their lives, compared to those who reside in the center. However, despite these differences, Haifa-bay subjects did not pay more attention to the logical quality of the policy plan, and they were affected by the symbols in the communication not less than those from the center. If anything, the results suggest that the symbols had a greater effect on Haifa-bay subjects. Furthermore, priming subjects to think about the personal relevance of environmental policy, and of the consequences of air-pollution in particular, also did not moderate the effects of the symbols and information in the communications.

How can we explain the finding that the positive effect of the real symbols was stronger among Haifa-Bay residents? Given that this pattern is also evident for the recycling policy, it is unlikely that it is linked to differences in perceived personal relevance. Rather, they can be explained by the lower baseline of Haifa-Bay subjects' trust in both policies, presumably due to their negative prior attitudes towards EPM. Accordingly, it could be the case that symbolic elements are more effective in shifting people from low to medium levels of trust (i.e. as in the case of Haifa-Bay subjects), than shifting them from medium to high levels (i.e. as in the case of center subjects), which results in a ceiling effect. Stated otherwise, symbols may be more effective in reducing citizens' distrust, as opposed to increasing their trust.[^footnote_ceiling_effect] Another related explanation is that the symbols distracted the subjects and decreased their inclination to think about this topic. As a result, they caused Haifa-bay subjects to disregard their prior beliefs about EPM and its actions regarding the air-pollution.[^footnote_mediation_elaboration]:

[^footnote_ceiling_effect]: This tentative explanation is supported by the fact that the interaction remains intact when controlling for demographic differences between the areas. It is also supported by the comparison of the percentages of subjects who reported very low levels of trust in the policies. This comparison shows a pattern of decreased marginal reduction. In the Haifa-Bay, the percentage of those who reported a trust score of less than 2 in the air-pollution policy decreased by from `r ((filter(SVIVA2_01_haifa,SYMBOL==0,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==0)%>%nrow())*100)%>%round(1)`% in the no symbols group to `r ((filter(SVIVA2_01_haifa,SYMBOL==2,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==2)%>%nrow())*100)%>%round(1)`%. In the center, only `r ((filter(SVIVA2_01_center,SYMBOL==0,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==0)%>%nrow())*100)%>%round(1)`% of the no symbols group reported a trust score lower than 2 in the first place, and this percentage dropped to `r ((filter(SVIVA2_01_center,SYMBOL==2,TRUST_air_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==2)%>%nrow())*100)%>%round(1)`% among the real symbols group. With regard to the recycling policy, the percentage decreased from `r ((filter(SVIVA2_01_haifa,SYMBOL==0,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==0)%>%nrow())*100)%>%round(1)`% to `r ((filter(SVIVA2_01_haifa,SYMBOL==2,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_haifa,SYMBOL==2)%>%nrow())*100)%>%round(1)`% in the Haifa-bay and from `r ((filter(SVIVA2_01_center,SYMBOL==0,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==0)%>%nrow())*100)%>%round(1)`% to `r ((filter(SVIVA2_01_center,SYMBOL==2,TRUST_waste_INDEX<2) %>% nrow())/(filter(SVIVA2_01_center,SYMBOL==2)%>%nrow())*100)%>%round(1)`% in the center.

[^footnote_mediation_elaboration]: This explanation is further supported by the survey data. The real symbols had a negative effect on the elaboration of these policies,  as measured by reaction time and memory test. The decrease in elaboration mediates the effect of symbols on trust. These analyses are presented in the supplentary appendix.

How can we reconcile the findings of this study with the proposed theory? One plausible explanation is that despite the relatively high involvement of Haifa-bay residents, they were still not motivated enough to rely only on the central route and disregard the peripheral cues. If this explanation holds, it entails that citizens may resist the emotional effect of symbols only in rare circumstances, which are characterized by extremely high levels of personal relevance. Indeed, Petty and Cacioppo have admitted that the high and low personal relevance conditions generated in their experimental settings were deliberately created for theory testing purposes, and can be rarely found in the world outside the laboratory [@petty_1986, p. 206]. The findings in this study are also consistent with the findings of @alonbarkat_2017, where they found that citizens who recently personally experienced prolonged blackouts (and presumably perceived electricity services as personally relevant) were nonetheless significantly affected by a symbol of the government-owned electricity company.

The results of this study, if generalizable, have significant normative implications. They challenge the optimistic view of government direct communications with citizens as a means to inform citizens about government organizations actions and promote transparency and accountability. Instead, they suggest that more often than not, citizens rely on symbols and other peripheral cues in public communications, including when these communications concern issues that have direct implications on thrir own lives. This inclination is in tension with the assumption that citizens should form their opinions about their government and about public affairs, based on their rational thinking [@dahl_1973; @soroka_2010]. It also buttresses the concern that government organizations will strategically use symbolic communications to manipulate public opinion, and cause citizens to view them more favorably regardless (or even in contrast) to their actual capacities and achievements.

Notwithstanding these conclusions, the limitations of this study should be acknowledged. The study focuses, empirically, on one specific case. Future studies may put these findings to further test by examining additional cases of public organizations and policy areas, as well as different forms of communications. Additionally, the comparison between the two areas is imperfect, insofar as the differences in personal relevance are confounded with differences in prior beliefs about the subject, which I cannot effectively control. A comparison between two groups that have similar prior beliefs would be more reliable. Finally, one can question the validity of relevance manipulation, which has not yielded significant effect on subjects' subjective reporting on the perceived personal relevance of the communications. Still, the main results are robust to different model specifications, and operationalizations of personal relevance and message elaboration, which substantiate their validity.

##Appendix A: Information manipulation

**Air-pollution policy** (Reducing the air-pollution in the Haifa-Bay)

EPM is working to reduce the air-pollution in the Haifa-Bay through:

*Strong*

* **Increasing the supervision of the factories in Haifa-Bay**.
In order to reduce the emissions of pollutants from the factories in the industrial area, the Ministry is working to increase the supervision of the factories in several ways. The ministry will increase the number of inspection visits, expand the use of unannounced inspections, and increase the amount of fines for pollutants.

* **Reduction of smoke emissions from vehicles in Haifa Bay**.
The Ministry is taking a number of actions to reduce the amount of air-pollution from vehicles in the area. The ministry promotes the installation of particle filters in hundreds of vehicles in the area, and promotes the operation of a "clean-air area" that restricts the entry of polluting vehicles.

*Weak*

* **Decreasing the supervision of the factories in Haifa-Bay**.
The Ministry is preparing for a change in the method of supervision of the factories. The ministry will reduce the number of inspection visits and unannounced inspections, rely on factories' self-reports on pollutant emissions, and reduce the amount of fines. The change is coordinated with members of the industry, which are committed to reducing the pollution, and it will assist to build mutual respect between them and the ministry.

* **Raising citizens' awareness of the efforts to improve air quality**.
In recent years, the Ministry has taken steps to reduce air pollution in Haifa-Bay, which have been partially successful. The Ministry is working to raise awareness of its efforts and to emphasize their success in order to improve its image among the residents. To this end, the ministry has increased its investment in public relations and publicity in this subject.

<br>

**Recycling policy** (Reducing waste and increasing recycling)

EPM is working to reduce the amount of waste and increase its recycling through:

*Strong*

* **Packaging recycling - Increased supervision of manufacturers**.
In order to reduce the amount of waste generated from packaging and to encourage recycling, the Ministry implements the Packaging act, which imposes on manufacturers and importers the responsibility for recycling the packaging. The Ministry will work to increase enforcement by virtue of this act, and will enforce it with regard to additional manufacturers and importers.

* **Reducing the usage of disposable bags**.
In order to reduce the use of disposable plastic bags, the Ministry enforces the Bags act, which sets a compulsory charge on bags sold in large retail chains and imposes reporting obligations on the chains. The Ministry works to implement the act, and encourages the transition to reusable grocery bags.

*Weak*

* **Packaging recycling - Decreased supervision of manufacturers**.
The Ministry is preparing to a change in the method of supervision of the manufacturers, who are responsible for the recycling of product packaging by virtue of the act The ministry will reduce the number of inspections, rely on manufacturers' self-reports regarding their compliance with the requirements of the act, and reduce the amount of fines. The change is coordinated with members of the industry, which are committed to the issue, and it will assist to strengthen mutual respect between them and the ministry.

* **Reducing the usage of disposable bags**.
In recent years, the Ministry has taken steps to encourage the reduction of waste, which have been partially successful. The Ministry works to raise awareness of the actions taken and to highlight their success in order to improve its image among the public. To this end, the ministry has increased its investment in public relations and publicity in this subject.

<br>

##Appendix B: Symbols manipulation
![](D:/UNIVERSITY/R/SVIVA/experiment 2 01-2018/myfigures/symbols_manipulation.png)

<font size="0.5">

*Note: The policy plans communications presented here are the strong policies, in the version adjusted for mobile interface.*

</font>

<br>

<br>

#References